---
title: "Load data and cleaning"
author: "Justin Williams"
date: "5/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(vroom)
library(sf)
library(tigris)
options(tigris_use_cache = TRUE)
library(leaflet)
library(htmlwidgets)
library(BAMMtools)
library(janitor)
library(scales)
library(RSocrata)
library(dotenv)
```

Set working directory

```{r setwd}
#setwd("/Users/justinwilliams/projects/nyc_covid_tracker")
setwd("C:/Users/jtws/OneDrive - IKEA/Projects/nyc_covid_tracker")
```

## Load in Covid data.

Download master repo from NYC DoH Github.

```{r doh-github}
# download data
download.file(url = 
    "https://github.com/nychealth/coronavirus-data/archive/refs/heads/master.zip",
    destfile = "covid_data_master.zip")

# unzip file
unzip(zipfile = "covid_data_master.zip")
```

### NYC DOHMH Github data

Sort through the following folders to see what data will actually be useful for this analysis.

**trends** 
*Weekly Data by Zip Code*
Would like choropleth maps with reactive selection box for each of the following. Also would like a sliderinput for week ending date. There could be a drop down within the tab
- `percentpositve-by-modzcta.csv` - percent of people who tested positive with a molecular test stratified by week.
- `caserate-by-modzcta.csv` this has rate of cases per 100,000 people by week and modzcta so could help for case rate, but not count. Could have graduated symbology be case rate or choropleth colors.
- `test-rate-by-modzcta.csv` - rate of molecular testing per 100,000 people stratified by week.

*Summary Numbers*
The idea is these number would be in a sidebar the entire time.
Cases
Deaths
Date
New Cases bar graph
Cumulative cases line chart
- `cases-by-day.csv` 
  - would be good for bar graph detailing new cases.
  - also works for cumulative line graph
- `summary.csv` - the has total case counts and deaths

**latest**
*Latest Data by Zip Code*
Could be used for Percent Positive (7-day), Test Rate (7-day), Hosp (28-day) and Death Rate (28-day) choropleth maps. Drop down within tab?
- `last7days-by-modzcta.csv` - has lat long 7 day pp rate along with people_tested, people_positive, median_daily_test_rate, adequately_tested (boolean) and daterange. 
- `hosp_death_last28days-by-modzcta.csv` - hosp/death count 28 day and hosp/death rate 28 day, daterange

### Load in data

Load in some .csv data

```{r load-data}
# for weekly maps tab
caserate <- vroom("./coronavirus-data-master/trends/caserate-by-modzcta.csv")
testrate <- vroom("./coronavirus-data-master/trends/testrate-by-modzcta.csv")
percentpos <- vroom("./coronavirus-data-master/trends/percentpositive-by-modzcta.csv")

# for summary sidebar
casesbyday <- vroom("./coronavirus-data-master/trends/cases-by-day.csv")
summary_totals <- vroom("./coronavirus-data-master/totals/summary.csv",
                        delim = ",")

# for latest tab
last7days <- vroom("./coronavirus-data-master/latest/last7days-by-modzcta.csv")
last28days <- vroom("./coronavirus-data-master/latest/hosp_death_last28days-by-modzcta.csv")
```

Load in MODZCTA geography from NYC Open Data.

```{r MODACTA-geography, message=FALSE}
# read from nyc open data
# this has population and zcta
modzcta <-st_read("https://data.cityofnewyork.us/resource/pri4-ifjk.geojson")
```

### Clean Data
#### Weekly Data

Delete columns and pivot longer.

```{r drop-col-pivot}
# function to drop columns and pivot longer
pivot_covid_long <- function(df, names_prefix, values_to) {
  df %>% 
    select(-c(2:7)) %>% 
    pivot_longer(2:178, names_to = "modzcta",
                 names_prefix = names_prefix,
                 values_to = values_to)
}

# apply to dfs
caserates_long <- pivot_covid_long(caserate, names_prefix = "CASERATE_", 
                              values_to = "caserate")
testrates_long <- pivot_covid_long(testrate, names_prefix = "TESTRATE_",
                              values_to = "testrate")
percpos_long <- pivot_covid_long(percentpos, names_prefix = "PCTPOS_",
                            values_to = "percpos")
```

Join df's

```{r join-format}
# subset df with neigborhood names first to join with neighborhood names
modzcta_name <-  last7days[,c("modzcta","modzcta_name")]
modzcta_name$modzcta <- as.character(modzcta_name$modzcta)

# join all df's
all <-caserates_long %>% 
  left_join(percpos_long, by = c("week_ending", "modzcta")) %>% 
  left_join(testrates_long, by = c("week_ending", "modzcta")) %>% 
  left_join(modzcta_name)
```

Add in spatial data.

```{r join-spatial-data, warning=FALSE}
# join with modzcta
all_sf <-geo_join(modzcta, all, 
                  "modzcta","modzcta",
                  how = "inner")

# convert week ending column to date
all_sf$week_ending <- as.Date(all_sf$week_ending, format = "%m/%d/%Y")

# sort by date to have most current date first
all_sf <- all_sf %>% 
  arrange(desc(week_ending))
```

**latest**
*Latest Data by Zip Code*
Could be used for Percent Positive (7-day), Test Rate (7-day), Hosp (28-day) and Death Rate (28-day) choropleth maps. Drop down within tab?
- `last7days-by-modzcta.csv` - has lat long 7 day pp rate along with people_tested, people_positive, median_daily_test_rate, adequately_tested (boolean) and daterange. 
- `hosp_death_last28days-by-modzcta.csv` - hosp/death count 28 day and hosp/death rate 28 day, daterange

#### Latest Data by Zip Code

Let's join this to the spatial data so we can map it. Data is clean, needs no other modifications.

```{r last-7-join-modzcta, warning=FALSE}
# convert modzcta to char
last7days$modzcta <- as.character(last7days$modzcta)
#drop label
last7days <- last7days %>% 
  select(-c("label"))

# join with spatial df
all_sf_7 <- geo_join(modzcta, last7days, 
              "modzcta","modzcta",
              how = "inner")
```

Let's also join 28 day data with spatial data.

```{r last-28-day-spatial, warning=FALSE}
# convert modzcta to char
last28days$modzcta <- as.character(last28days$modzcta)
#drop label
last28days <- last28days %>% 
  select(-c("label"))

# join with spatial df
all_sf_28 <- geo_join(modzcta, last28days, 
              "modzcta","modzcta",
              how = "inner")
```

#### Summary data
*Summary Numbers*
The idea is these number would be in a sidebar the entire time.
Cases
Deaths
Date
New Cases bar graph
Cumulative cases line chart
- `cases-by-day.csv` 
  - would be good for bar graph detailing new cases.
  - also works for cumulative line graph
- `summary.csv` - the has total case counts and deaths

Summary numbers

```{r summary}
#case count
summary_totals[summary_totals$MEASURE == "NYC_TOTAL_CASE_COUNT",]$NUMBER_OF_NYC_RESIDENTS

# death count
summary_totals[summary_totals$MEASURE == "NYC_TOTAL_DEATH_COUNT",]$NUMBER_OF_NYC_RESIDENTS

# date
summary_totals[summary_totals$MEASURE == "DATE_UPDATED",]$NUMBER_OF_NYC_RESIDENTS
```

New Cases

```{r new-cases}
# convert date column to date format
casesbyday$date_of_interest <- as.Date(casesbyday$date_of_interest, format = "%m/%d/%Y")

# get past 6 months
six_months <- max(casesbyday$date_of_interest) - 180

# segment df for only past 6 months
casesbyday_6_mon <- casesbyday[casesbyday$date_of_interest >= six_months,]

# subset df
casesbyday_6_mon <-  casesbyday_6_mon %>% 
  rename(date  = date_of_interest,
         case_count = CASE_COUNT) %>% 
  select(c("date", "case_count"))
```

Cases past 6 months plot.

```{r plot-cases-past-6-months}
casesbyday_6_mon %>% 
  ggplot(aes(date, case_count)) +
    geom_bar(stat = "identity") +
    scale_y_continuous(labels = number_format(scale = .0001,
                                              suffix = "k")) +
    scale_x_date(date_breaks = "1 month",
                 date_labels = "%b-'%y") +
    labs(y = "", x = "", 
         title = "Case Count by Day Last 6 Months") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 60, hjust = 1),
          plot.title.position = "plot",
          plot.title = element_text(hjust = 0.5, size = 18),
          axis.text = element_text(size = 10))
```
Cases since inception

```{r cases-since--inception}
casesbyday %>% 
  ggplot(aes(date_of_interest, CASE_COUNT)) +
    geom_bar(stat = "identity") +
    scale_y_continuous(labels = number_format(scale = .0001,
                                              suffix = "k")) +
    scale_x_date(date_breaks = "3 month",
                 date_labels = "%b-'%y") +
    labs(y = "", x = "", 
         title = "Case Count by Day Last 6 Months") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 60, hjust = 1),
          plot.title.position = "plot",
          plot.title = element_text(hjust = 0.5, size = 18),
          axis.text = element_text(size = 10))
```


Cumulative cases
```{r cumulative-cases}
# create cumsum column in casesbyday
casesbyday_cum <- casesbyday %>% 
  mutate(case_cum = cumsum(CASE_COUNT)) %>% 
  rename(date = date_of_interest,
         case_count = CASE_COUNT) %>% 
  select(c("date", "case_count", "case_cum"))
```

Cumulative cases line plot

```{r line-plot-cum-cases}
casesbyday_cum %>% 
  ggplot(aes(date, case_cum)) +
    geom_line(lwd = 2) +
    geom_point(
      aes(x = as.Date('2022-01-01'), y = 1000000), size = 3, color = 'Red') +
    scale_y_continuous(labels = number_format(scale = .0001,
                                              suffix = "k")) +
    scale_x_date(date_breaks = "3 month",
                 date_labels = "%b-'%y") +
    labs(y = "", x = "", 
         title = "Cumulative Cases") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 60, hjust = 1),
          plot.title.position = "plot",
          plot.title = element_text(hjust = 0.5, size = 18),
          axis.text = element_text(size = 10))

```


### Leaflet map 

#### Weekly data

Write code for leaflet map prior to creating Shiny App.
Redo this for just one week-ending, its too much data and takes forever to render.

```{r leaflet-map}
# segment for week ending 2022-01-01
all_sf_jan_01_22 <- all_sf[all_sf$week_ending == "2022-01-01",]

# set labels
labels <- sprintf(
  "<h3>%s</h3><br/><strong>Zip Code: </strong>%s<br/><strong>Cases per 100,000 people: </strong>%g ",
  all_sf_jan_01_22$modzcta_name,
  all_sf_jan_01_22$label, round(all_sf_jan_01_22$caserate)) %>% 
  lapply(htmltools::HTML)

# set bins
bins <- getJenksBreaks(all_sf_jan_01_22$caserate, 6) # get natural breaks
bins <- map(.x = bins, round) %>% 
  unlist() # round

# set color palette
pal <- colorBin(palette = "PuBu",
                bins = bins, 
                domain = all_sf_jan_01_22$caserate)

# plot map
leaflet_map <- all_sf_jan_01_22 %>% 
   st_transform(4326) %>% 
   leaflet() %>%
   setView(lng = -74.0060,lat = 40.7128,zoom = 10) %>% 
   addProviderTiles(providers$CartoDB.Positron) %>% 
   addPolygons(label = labels,
               weight = 0.25,
               color = "white",
               smoothFactor = 0.5,
               opacity = 1,
               fillOpacity = 0.7,
               fillColor = pal(all_sf_jan_01_22$caserate),
               highlightOptions = highlightOptions(weight = 1,
                                                   color = "#666",
                                                   fillOpacity = 0.7,
                                                   bringToFront = T)) %>% 
  addLegend("bottomright",
            pal = pal,
            values = ~caserate,
            title = "Cases per 100,000",
            opacity = 0.7)

mapview::mapshot(leaflet_map,
                 file = "./images/cases_draft_map.png")
leaflet_map

all_sf_jan_01_22
```

#### Latest Data maps

Create choropleth maps with latest data.

```{r 7-day-map}
# set labels
labels <- sprintf(
  "<h3>%s</h3><br/><strong>Zip Code: </strong>%s<br/><strong>7-day percent positive: </strong>%g%% ",
  all_sf_7$modzcta_name,
  all_sf_7$label, all_sf_7$percentpositivity_7day) %>% 
  lapply(htmltools::HTML)

# set bins
bins <- getJenksBreaks(all_sf_7$percentpositivity_7day, 6) # get natural breaks
bins <- map(.x = bins, round) %>% 
  unlist() # round

# set color palette
pal <- colorBin(palette = "PuBu",
                bins = bins, 
                domain = all_sf_7$percentpositivity_7day)

# plot map
leaflet_map <- all_sf_7 %>% 
   st_transform(4326) %>% 
   leaflet() %>%
   setView(lng = -74.0060,lat = 40.7128,zoom = 10) %>% 
   addProviderTiles(providers$CartoDB.Positron) %>% 
   addPolygons(label = labels,
               weight = 0.25,
               color = "white",
               smoothFactor = 0.5,
               opacity = 1,
               fillOpacity = 0.7,
               fillColor = pal(all_sf_7$percentpositivity_7day),
               highlightOptions = highlightOptions(weight = 1,
                                                   color = "#666",
                                                   fillOpacity = 0.7,
                                                   bringToFront = T)) %>% 
  addLegend("bottomright",
            pal = pal,
            values = ~caserate,
            title = "7-day percent positive",
            opacity = 0.7,
            labFormat = labelFormat(suffix = "%"))


leaflet_map
```

28-day map

```{r 28-day-hosp-map}
# set labels
labels <- sprintf(
  "<h3>%s</h3><br/><strong>Zip Code: </strong>%s<br><strong>Hospitalization rate per 100,000 people: </strong>%g <br><strong>Hospitalization Count: </strong>%g",
  all_sf_28$modzcta_name,
  all_sf_28$label, all_sf_28$hospitalization_rate_28day,
  all_sf_28$hospitalization_count_28day) %>% 
  lapply(htmltools::HTML)

# set bins
bins <- getJenksBreaks(all_sf_28$hospitalization_rate_28day, 5) # get natural breaks
bins <- map(.x = bins, round) %>% 
  unlist() # round

# set color palette
pal <- colorBin(palette = "PuBu",
                bins = bins, 
                domain = all_sf_28$hospitalization_rate_28day)

# plot map
leaflet_map <- all_sf_7 %>% 
   st_transform(4326) %>% 
   leaflet() %>%
   setView(lng = -74.0060,lat = 40.7128,zoom = 10) %>% 
   addProviderTiles(providers$CartoDB.Positron) %>% 
   addPolygons(label = labels,
               weight = 0.25,
               color = "white",
               smoothFactor = 0.5,
               opacity = 1,
               fillOpacity = 0.7,
               fillColor = pal(all_sf_28$hospitalization_rate_28day),
               highlightOptions = highlightOptions(weight = 1,
                                                   color = "#666",
                                                   fillOpacity = 0.7,
                                                   bringToFront = T)) %>% 
  addLegend("bottomright",
            pal = pal,
            values = ~caserate,
            title = "Hospitalizations per 100,000 people",
            opacity = 0.7)


leaflet_map
```

### Export data 

Export data frames for use in Shiny App.
First get list

```{r df-list}
Filter(function(x) is(x, "data.frame"), mget(ls()))
```

Create list of df to export then loop through and export to app folder.

```{r export-data}
# list of df
df_export <- c("all_sf",
               "all_sf_28",
               "all_sf_7",
               "casesbyday_6_mon",
               "casesbyday_cum",
               "summary_totals")

# save for use in shiny app
for (i in 1:length(df_export)) {
  savefile <- paste0(paste0("nyc_covid_tracker_app/", df_export[i], ".Rds"))
  saveRDS(get(df_export[i]), file = savefile)
  print(paste("Dataframe Saved: ", df_export[i]))
}
```

